Gravitational wave sources in our Galactic backyard: Predictions for BH and NS binaries in LISA

### 2021-07-08 paper/tex/sections/comparisons.tex
Updates from Overleaf

In Figure~\ref{fig:previous_studies}, we compare similar previous studies that have investigated the population of stellar mass BHBHs, BHNSs and NSNSs that are detectable with LISA. The table details the expected detection rates predicted by each paper as well as their assumptions regarding their Milky Way galaxy model, binary population synthesis simulation and LISA mission specifications. This table not include the number papers on the LISA WDWD population as we do not make predictions for these DCOs. We additionally do not include papers that don't use population synthesis or do not simulate sources in the Galactic plane \citep[e.g.][]{Andrews+2020, Kremer+2018} such that we only compare with papers with similar methodology and focus.

\citet{Nelemans+2001} were the first to investigate the population of LISA detectable stellar mass double compact objects. Their predictions of the number of detectable DCOs are very different from our own, where we find a significantly higher detection rate for BHBHs and BHNSs, as well as a slightly lower rate for NSNSs. However, since the publication of their paper, there have been many changes both to the specifications of LISA (such as the mission length and SNR threshold for detection) and our understanding of massive star evolution, which strongly affect the expected detections rates.

\citet{Belczynski+2010} built upon this work by using a different population synthesis code with two model variations and a multi-component model for the Milky Way. They found a similar number of BHBH and BHNS detection to \citet{Nelemans+2001}, but a lower number of NSNS detections. The low total detection rate for all DCOs in this paper is unsurprising given the relatively high SNR threshold of 10 and short mission length of 1 year. The reduced mission length means that the source signal has much less time to accumulate, whilst also fewer WDWDs can be resolved in this time, leading to a weaker signal and an increased Galactic confusion noise relative to our work.

More recently, \citet{Lamberts+2018} presented a new approach to the problem by using the FIRE simulation \citep{Hopkins+2014} to distribute their sources rather than an analytical model of the Milky Way, thus being the first paper in this area to incorporate metallicity dependence into their Milky Way model. \citet{Sesana+2020} followed up on this paper using the same simulated BHBH population and presented updated results for the number of expected BHBH detections. They find significantly fewer BHBHs than our fiducial model despite using the same SNR threshold and LISA mission length.

The discrepancy between their results and ours could be a result of different implicit assumptions in their population synthesis code, however we expect that BSE should be similar to COMPAS since it provided the basis of the COMPAS code. Another major difference is that, unlike this work, they assume that all binaries are circular for the purpose of detection in LISA, which could result in a lower number of detections by missing eccentric binaries that appear as weaker signals when assumed to be circular. We also improve upon this work by using a larger number of metallicity bins since a low number of metallicity bins can produce artificial features in the mass distribution of DCOs and possibly affect the detection rate.

>>>>

In Figure~\ref{fig:previous_studies}, we compare our results to similar previous studies that investigate the population of stellar mass BHBHs, BHNSs and NSNSs that are detectable with LISA. The table details the expected detection rates predicted by each paper as well as their assumptions regarding their Milky Way galaxy model, binary population synthesis simulation and LISA mission specifications. This table does not include the number papers on the LISA WDWD population as we do not make predictions for these DCOs. We additionally do not include papers that don't use population synthesis or do not simulate sources in the Galactic plane \citep[e.g.][]{Andrews+2020, Kremer+2018} such that we only compare with papers with similar methodology and focus.

\citet{Nelemans+2001} were the first to investigate the population of LISA-detectable stellar mass double compact objects. Their predictions of the number of detectable DCOs are very different from our own, where  We find a significantly higher detection rate for BHBHs and BHNSs, as well as a slightly lower rate for NSNSs. However, Since the publication of their paper, there have been many changes both to the specifications of LISA (such as the mission length and SNR threshold for detection) and our understanding of massive star evolution, which strongly affect the expected detections rates.

\citet{Belczynski+2010} built upon the work of \citet{Nelemans+2001}, by using a different population synthesis code with two model variations and a multi-component model for the Milky Way. They found a similar number of BHBH and BHNS detection to \citet{Nelemans+2001}, but a lower number of NSNS detections. The low total detection rate for all DCOs in this paper is unsurprising given the relatively high SNR threshold of 10 and short mission length of 1 year. The reduced mission length means that the source signal has much less time to accumulate, whilst also fewer WDWDs can be resolved in this time, leading to a weaker signal and an increased Galactic confusion noise relative to our work.

More recently, \citet{Lamberts+2018} presented a new approach to the problem by using the FIRE simulation \citep{Hopkins+2014} to distribute their sources rather than an analytical model of the Milky Way, thus being the first paper in this area to incorporate metallicity dependence into their Milky Way model. \citet{Sesana+2020} followed up on this paper using the same simulated BHBH population and presented updated results for the number of expected BHBH detections. They find significantly fewer BHBHs than our fiducial model despite using the same SNR threshold and LISA mission length.
%
The discrepancy between their results and ours could be a result of different implicit assumptions in their population synthesis code. However, we expect that BSE should be similar to COMPAS since it provided the basis of the COMPAS code. Another major difference is that, unlike this work, they assume that all binaries are circular for the purpose of detection in LISA, which could result in a lower number of detections by missing eccentric binaries that appear as weaker signals when assumed to be circular. We also improve upon this work by using a larger number of metallicity bins since a low number of metallicity bins can produce artificial features in the mass distribution of DCOs and possibly affect the detection rate.

### 2021-07-08 paper/tex/sections/methods.tex
Updates from Overleaf

The star formation history of the Milky Way bulge (which we assume here to be dominated by the central bar) has many uncertainties due to (1) sizeable age measurement uncertainties at large ages in observational studies, (2) complex selection processes affecting the observed age distributions, and (3) formation mechanisms still under debate. But the central bar was shown to contain stars with an age range of 6-12 Gyr \citep[e.g.,][]{Bovy+2019}, where the younger tail of ages might come from the growth of the Galactic bar. To model the bar's age distribution more realistically than in previous studies (assuming an old bulge coming from a single starburst), we choose to adopt a star formation history using a $\beta(2,3)$ distribution, shifted and scaled such that stars are only formed in the range $[6, 12] \unit{Gyr}$. We show these distributions in the first panel in the left half of Fig.~\ref{fig:galaxy_schematic}.

>>>>

The star formation history of the Milky Way bulge (which we assume here to be dominated by the central bar) has many uncertainties due to (1) sizeable age measurement uncertainties at large ages in observational studies, (2) complex selection processes affecting the observed age distributions, and (3) formation mechanisms still under debate. But the central bar was shown to contain stars with an extended age range, with most observed stars between 6 and 12 Gyr \citep[e.g.,][]{Bovy+2019}, and a younger tail of ages that could come from the subsequent secular growth of the Galactic bar. To model the bar's age distribution more realistically than in previous studies (which assume an old bulge coming from a single starburst), we choose to adopt a more extended star formation history using a $\beta(2,3)$ distribution, shifted and scaled such that stars are only formed in the range $[6, 12] \unit{Gyr}$. We show these distributions in the first panel in the left half of Fig.~\ref{fig:galaxy_schematic}.
